---
title: "Practical Machine Learning Course Project"
author: "A. Sachin Vianney"
date: "October 7, 2018"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

# Executive Summary
Three different Machine learning algorithm are applied to the training set and is used to predict using the test set with varied accuracies.We can infer from the analysis that Random Forest is the best model and is applied to the test set.

We first need to include all the essential libraries/packages.

```{r chunk1}
library(ggplot2)
library(rattle)
library(caret)
library(rpart)
library(randomForest)
```

Initailly, we need to download the test and training datasets from the URLs given to us. We set our working directory and proceed with the download followed by reading of the datasets.

```{r chunk2}
#INPUTTING THE DATA

training <- read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"),header=TRUE)
test <- read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"),header=TRUE)
```

Displaying the raw uncleaned data

```{r chunk3}
#DISPLAYING THE RAW DATA

str(training)

```

As we can see, there are a lot of NA values, we need to clean this data in order to make our analysis easier and more meaningful.

```{r chunk4}
#DATA CLEANING

na_count <- sapply(test, function(y) sum((is.na(y))))
na <- na_count[na_count == 20]
good <- names(na)
training <- training[,!(names(training) %in% good)]
test <- test[,!(names(test) %in% good)]
good2 <- c('user_name','raw_timestamp_part_1', 'raw_timestamp_part_2', 'cvtd_timestamp', 'new_window', 'num_window', 'X')
training <- training[,!(names(training) %in% good2)]
test <- test[,!(names(test) %in% good2)]

```

Displaying the clean data, we can see a reduction in the number of rows

```{r chunk5}
#DISPLAYING THE CLEAN DATA

str(training)

```

# Machine Learning Algorithms

For this analysis, we will be using three algorithmic approaches
The Three algorithms applied are:
1)Decision Tree
2)Random Forest
3)Generalized Boosted Regression

```{r chunk6}
#MACHINE LEARNING ALGORITHMS IMPLEMENTATION

set.seed(49)
part <- createDataPartition(training$classe, p = 0.75, list = FALSE)
training1 <- training[part,]
test1 <- training[-part,]

```

I have used the cross-validation technique in order to improve efficiency and limit overfitting.I have used 10 folds.

```{r chunk7}
tr_cont <- trainControl(method="cv", number=10)
```

# Decision Trees

```{r chunk8}
model1 <- train(classe~., data=training1, method="rpart", trControl=tr_cont)
fancyRpartPlot(model1$finalModel)
prediction1 <- predict(model1,newdata=test1)
conf_mat1 <- confusionMatrix(test1$classe,prediction1)
conf_mat1$table
conf_mat1$overall[1]

```

# Random Forest

```{r chunk9}
model2 <- train(classe~., data=training1, method="rf",trControl=tr_cont,verbose = FALSE)
print(model2)
prediction2 <- predict(model2,newdata=test1)
conf_mat2 <- confusionMatrix(test1$classe,prediction2)
conf_mat2$table
conf_mat2$overall[1]

```

# Generalized Boosted Regression

```{r chunk10}
model3 <- train(classe~., data=training1, method="gbm", trControl=tr_cont, verbose=FALSE)
print(model3)
prediction3 <- predict(model3,newdata=test1)
conf_mat3 <- confusionMatrix(test1$classe,prediction3)
conf_mat3$table
conf_mat3$overall[1]

```

Since the Random Forests gave us a prediction with the highest accuracy, we will go with model 2 for the final prediction.

# The Final Result/Prediction

```{r chunk11}
final <- predict(model2,newdata= test)
final

```